\section{Dataset}

\subsection{Source and Overview}

We use the \textit{USA House Prices} dataset published on Kaggle~\footnote{\url{https://www.kaggle.com/datasets/fratzcan/usa-house-prices/data}},
which contains 4,140 residential property sales records from Washington State.
The raw dataset comprises 18 columns covering structural attributes, location
information, and transaction metadata.
Preprocessing removes records with a non-positive price and the most extreme
price outliers (the top $0.1\%$ by price), leaving \textbf{4,086} records; this
caps only a handful of pathological sales while deliberately retaining realistic
noise.
We then remove non-informative fields (\texttt{date}, \texttt{street},
\texttt{statezip}, \texttt{country}) and apply the feature engineering described
in Section~\ref{sec:method}, yielding \textbf{17} model-ready input features.
All reported numbers come from 5-fold cross-validation over these 4,086 records
(about 3,269 training / 817 validation rows per fold); we do not use a single
fixed train/test split.

\subsection{Prediction Target}

The single reported prediction target is \textbf{\texttt{price}}, the raw
transaction price in USD. As is typical for residential price prediction, raw
price is strongly correlated with living area (\texttt{sqft\_living}): larger
homes tend to be more expensive simply by virtue of their size, which can cause
a model to over-rely on size-related features.

To counteract this, we additionally derive an auxiliary quantity,
\texttt{price\_per\_sqft}:
\begin{equation}
  \texttt{price\_per\_sqft} = \frac{\texttt{price}}{\texttt{sqft\_living}}.
\end{equation}
Normalising sale price by interior area---analogous to the ``price per square
metre'' convention in real estate valuation---decouples the size effect from the
price signal and tames price outliers (an extreme \emph{price} need not be an
extreme \emph{price-per-sqft}).
Crucially, \texttt{price\_per\_sqft} is \emph{not} reported as a separate
regression task. Instead it is used inside the stacking ensemble of
Section~\ref{sec:method}: a price-per-sqft model is trained alongside a direct
price model, its prediction is converted back to a price via $\times\,
\texttt{sqft\_living}$, and the two are blended. All four models (\textbf{MLP},
\textbf{DT}, \textbf{RF}, \textbf{mGBDT}) are trained and evaluated on the
\texttt{price} target, both as single models and as this stacked ensemble.

\subsection{Feature Description}

Table~\ref{tab:features} lists all 17 features used as model inputs after
preprocessing. Removed raw columns (\texttt{sqft\_lot}, \texttt{yr\_built},
\texttt{yr\_renovated}, \texttt{lat}, \texttt{long}) are replaced by their
engineered counterparts. The location fields \texttt{city} and \texttt{zipcode}
are encoded differently per experimental variant (ordinal codes vs.\ target
encoding), and the coordinate features \texttt{x,y,z} are present only in the
coordinate-aware variants; see Section~\ref{sec:method}.

\begin{table}[h]
\centering
\caption{Feature descriptions after preprocessing.}
\label{tab:features}
\begin{tabular}{llp{7.5cm}}
\toprule
\textbf{Feature} & \textbf{Type} & \textbf{Description} \\
\midrule
\texttt{bedrooms}        & Numeric   & Number of bedrooms \\
\texttt{bathrooms}       & Numeric   & Number of bathrooms (0.5-unit increments) \\
\texttt{sqft\_living}    & Numeric   & Interior living area (sq.\ ft.) \\
\texttt{floors}          & Numeric   & Number of floors \\
\texttt{waterfront}      & Binary    & Waterfront property (0 = No, 1 = Yes) \\
\texttt{view}            & Ordinal   & View quality score (0--4) \\
\texttt{condition}       & Ordinal   & Property condition score (1--5) \\
\texttt{sqft\_above}     & Numeric   & Above-ground floor area (sq.\ ft.) \\
\texttt{sqft\_basement}  & Numeric   & Basement area (sq.\ ft.) \\
\texttt{city}            & Categorical (encoded) & City name; ordinal code or target-encoded per variant \\
\texttt{x}               & Numeric   & 3-D Cartesian coordinate: $\cos(\text{lat})\cdot\cos(\text{lon})$ \\
\texttt{y}               & Numeric   & 3-D Cartesian coordinate: $\cos(\text{lat})\cdot\sin(\text{lon})$ \\
\texttt{z}               & Numeric   & 3-D Cartesian coordinate: $\sin(\text{lat})$ \\
\texttt{zipcode}         & Categorical (encoded) & ZIP code; ordinal code or target-encoded per variant \\
\texttt{log\_sqft\_lot}  & Numeric   & $\log(1+\texttt{sqft\_lot})$; log-transform of lot area to reduce skew \\
\texttt{was\_renovated}  & Binary    & Renovation history (0 = Never, 1 = Renovated) \\
\texttt{age}             & Numeric   & Building age at time of sale: sale year $-$ year built \\
\bottomrule
\end{tabular}
\end{table}
